---
title: "irisToolsHollarJ"
author: "Jadyn Hollar"
date: "2025-12-01"
output: html_document
---

Introduction

This R Markdown is for my R package irisToolsHollarJ. The intent of this document is to document how my functions operate on the standard iris data set. This should be a fairly straightforward read with some inline code to exemplify how my functions work.

This document will feature:

- loading of my package
- preparing of iris data
- summary of iris data
- scatterplot
- Tukey outlier function

Load the Package

To start, I load my package in addition to ggplot2 for plotting.

```{r}
library(devtools)
devtools::load_all(".") 
library(ggplot2)
```


Prepare the Package

For this I will be using my prepare_iris function that cleans and prepares the iris data. It also checks that it has all the required data  and that species is a factor. 
```{r}
clean_iris <- prepare_iris()
head(clean_iris)
```


Summary

This function creates a summary table for the datas by species of the sepal length using the iris_summary function. 
```{r}
summary_obj <- summarize_iris(clean_iris)

summary_obj
class(summary_obj)
```

Scatter plot

This function creates a scatter plot that plots the petal width and sepal length while categorizing them on color based on the species. To keep it simple, it uses the minimal theme. 
```{r}
plot_iris_scatter(
data = clean_iris,
xvar = "Petal.Width",
yvar = "Sepal.Length"
)
```

Tukey outlier

This function fixes bugs and also includes defensive checks by creating a matrix to run the function and determine if there are any significant outliers.
```{r}
set.seed(123)

test_matrix <- matrix(rnorm(50), nrow = 10)

test_matrix
tukey_multiple_safe(test_matrix)
```
